---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300,
  fig.width = 8,
  fig.height = 6
)
```


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
<!-- badges: end -->

## Overview

**bodymap** provides a simple and intuitive way to visualize body-related data in R. Whether you're mapping symptom frequencies, pain intensities, or regional distributions in medical research, bodymap makes it easy to create clear and professional anatomical visualizations using the familiar `ggplot2` syntax.

### Key Features

- 📊 **Ready-to-use body region dataset** with 14 anatomical regions
- 🎨 **Seamless ggplot2 integration** via `geom_sf()`
- 🔍 **Detailed anatomical coverage** including front/back views and spinal subdivisions
- 🏥 **Perfect for medical visualizations** such as symptom mapping, pain distribution, or injury patterns
- 📦 **Lightweight**: No functions, just clean spatial data

## Installation

### From GitHub

You can install the development version of bodymap from GitHub:
```{r, eval = FALSE}
# install.packages("pak")
pak::pak("benediktclaus/bodymap")
```

### From R-universe

Or install from my R-universe repository:
```{r, eval = FALSE}
# Enable repository
options(repos = c(
  benediktclaus = 'https://benediktclaus.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'
))

# Install bodymap
install.packages('bodymap')
```

## Quick Start
```{r example-basic}
library(bodymap)
library(ggplot2)

# Create a simple body map visualization
ggplot(bodymap) +
  geom_sf(fill = "lightblue", color = "black") +
  theme_void()
```

## Examples

### Distinguish Between Front and Back Views
```{r example-views}
library(dplyr)

bodymap |>
  mutate(view = ifelse(id >= 2000, "Back", "Front")) |>
  ggplot() +
  geom_sf(aes(fill = view), color = "black") +
  scale_fill_manual(values = c("Front" = "lightblue", "Back" = "lightcoral")) +
  labs(fill = "View") +
  theme_void()
```

### Highlight Spinal Regions
```{r example-spine}
library(rcartocolor)

ggplot(bodymap) +
  geom_sf(aes(fill = SUBREGION), color = "black") +
  scale_fill_carto_d(palette = "Safe", na.value = "grey90") +
  labs(fill = "Spinal Region") +
  theme_void()
```

### Map Pain Intensity Data
```{r example-pain}
# Create example pain data
pain_data <- data.frame(
  id = c(1010, 1030, 2051, 2052, 1040),
  intensity = c(0.7, 0.9, 0.6, 0.8, 0.4)
)

# Join with bodymap and visualize
bodymap |>
  left_join(pain_data, by = "id") |>
  ggplot() +
  geom_sf(aes(fill = intensity), color = "black") +
  scale_fill_carto_c(palette = "OrYel", na.value = "grey95", limits = c(0, 1)) +
  labs(fill = "Pain Intensity") +
  theme_void()
```

### Aggregate by Region Name

When you want to map the same value to all instances of a region (e.g., all "Back" subdivisions):
```{r example-aggregate}
# Symptom frequencies by region name
symptom_by_name <- data.frame(
  NAME = c("Head", "Chest", "Back"),
  frequency = c(0.8, 0.6, 0.5)
)

bodymap |>
  left_join(symptom_by_name, by = "NAME") |>
  ggplot() +
  geom_sf(aes(fill = frequency), color = "black") +
  scale_fill_carto_c(palette = "Purp", na.value = "grey95") +
  labs(fill = "Frequency") +
  theme_void()
```

## Dataset Structure

The `bodymap` dataset contains 14 body region polygons with the following variables:

- **id**: Unique identifier (10xx = front, 20xx = back)
- **NAME**: Body region name (Head, Face, Chest, Belly, Back, Upper/Lower Limbs, Genitals)
- **SUBREGION**: Spinal subdivisions (Cervical, Thoracic, Lumbar, Sacrum) where applicable
- **geometry**: Spatial multipolygon data (sf format)
```{r dataset-preview}
bodymap
```

## Use Cases

- 🏥 **Clinical research**: Visualize symptom distributions, pain patterns, or injury locations
- 📊 **Medical surveys**: Display body-related questionnaire results
- 🔬 **Epidemiology**: Map disease prevalence across body regions
- 💊 **Pharmacology**: Show drug efficacy or side effect locations
- 🏋️ **Sports medicine**: Illustrate injury patterns or training focus areas

## Getting Help

If you encounter a bug or have a feature request, please [open an issue](https://github.com/benediktclaus/bodymap/issues) on GitHub.

## License

This program is free software: you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the [GNU General Public License](https://www.gnu.org/licenses/gpl-3.0.html) for more details.

## Acknowledgments

Body regions were manually digitized for visualization purposes. This package builds on the excellent [sf](https://r-spatial.github.io/sf/) and [ggplot2](https://ggplot2.tidyverse.org/) packages.

---

**Author**: Benedikt Claus  
**GitHub**: [@benediktclaus](https://github.com/benediktclaus)
